\mytitle

### 2016-04-26 report/paper.tex
Fix errors

\begin{abstract}
One of the fundamental problems of distributed computing, is the ordering of events through all peers. Between different types of ordering, total ordering is of particular interest as it provides a powerful abstraction for building reliable distributed applications. unfortunately, existing algorithms can not provide reliability, scalability, resiliency and total ordering in one package. EpTO is a total order algorithm with probabilistic agreement that scales both in the number of processes and events. EpTO uses a probabilistic agreement to disseminate events through the system with high probability, and also provides integrity, total order and validity. We are going to implement EpTO using the NeEM library and show that EpTO is well-suited for large-scale dynamic distributed systems, and afterwards we will evaluate this algorithm by comparing it with a currently being used ordering algorithm.
\end{abstract}

\section{Introduction}
The ordering of events is one of the most fundamental problems in distributed systems and it has been studied over past few decades. Many researches have been done to create algorithms with different properties and trade-offs such as synchronization, agreement or state machine replication \autocites[]{defago2004total}[]{lamport1978time}. But due the fact that these properties are strong guarantees, the algorithms that implements them do not scale as expected because of the churn and failures in an non-ideal environment. Because of scalability problem of deterministic approaches, probabilistic protocols emerged. Nowadays, existing probabilistic protocols are highly scalable and resilient, but they do not provide total ordering.
\par
The problem with existing deterministic total ordering protocols, is that they need some sort of agreement between all peers in the system on the order of messages. This causes a massive amount of network traffic and overhead on the system.
Moreover, an agreement feature for an asynchronous system requires to
explicitly maintain a group and have access to a failure detector. Due to the faults and churn in large-scale distributed systems, failure detector turns into bottleneck of the system and thus, limits scalability of the algorithm.
\subsection{Contribution}
EpTO \autocite{matos2015epto} guarantees that all processes eventually agree on the set of received events with high probability and deliver these in total order to the application. The intuition behind EpTO is make events available in all processes quickly with high probability,
Once events are available in every peer, each peer deterministically order them using the timestamp of each event and breaking ties with the id of the broadcaster peer, and afterwards deliver them to the application accordingly.
\par
The main insight behind EpTO dissemination protocol is a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}. \textit{Balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin, completely random and independent from other balls.
\par
Using balls-and-bins approach we model processes an bins and events (or set of events) as balls. Base on this model we will calculate how many balls need to be \textit{thrown} such that each bin contains at least one ball with arbitrarily high probability. Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore number of messages sent in the system is low and uniform over all processes. Using these approaches, EpTO becomes highly scalable and resilient, having in mind that it also provide total order.

@@

\subsection{Total Order Algorithms}
A totally ordered set of events is a partially ordered set which satisfies one additional property:
\begin{itemize}
	\item Totality (trichotomy law): For any $a, b \in S$, either $a \leq b$  or $b \leq a$.
\end{itemize}
\todo[inline]{Explain advantages and usefulness of both systems}
\par
In other words, total order is an ordering that defines the exact order of every event in the system, But partial ordering only defines the order between certain key event that depend on each other. Partial order can be useful since it is less costly to implement but, in some cases where the order of all events is important, for example in order to have an exact image of the state of the system at all time, we have to use total order.

>>>>

\begin{abstract}
One of the fundamental problems of distributed computing, is the ordering of events through all peers. Among different types of ordering, total ordering is of particular interest as it provides a powerful abstraction for building reliable distributed applications. unfortunately, existing algorithms can not provide reliability, scalability, resiliency and total ordering in one package. EpTO is a total order algorithm with probabilistic agreement that scales both in the number of processes and events. EpTO uses a probabilistic agreement to disseminate events through the system with high probability, and also provides integrity, total order and validity. We are going to implement EpTO using the NeEM library and show that EpTO is well-suited for large-scale dynamic distributed systems. Afterwards we will evaluate EpTO by comparing it to the deterministic total order algorithm provided by JGroups.
\end{abstract}

\section{Introduction}
The ordering of events is one of the most fundamental problems in distributed systems and it has been studied over the past few decades. Many researches have been done to create algorithms with different properties and trade-offs such as synchronization, agreement or state machine replication \autocites[]{defago2004total}[]{lamport1978time}. Unfortunately, due the fact that these properties are strong guarantees, the algorithms that implement them do not scale as expected because of churn and failures in an non-ideal environment. Because of the scalability problem inherent to deterministic approaches, probabilistic protocols have emerged. Nowadays, existing probabilistic protocols are highly scalable and resilient, but they do not provide total ordering. \todo[inline]{Maybe add a paper to justify this claim}
\par
The problem with existing deterministic total ordering protocols, is that they need some sort of agreement between all peers in the system on the order of messages. This causes a massive amount of network traffic and overhead on the system.
Moreover, an agreement feature for an asynchronous system requires to
explicitly maintain a group and have access to a failure detector. Due to the faults and churn in large-scale distributed systems, failure detector turns into bottleneck of the system and thus, limits the scalability of the algorithm.
\subsection{Contribution}
EpTO \autocite{matos2015epto} guarantees that all processes eventually agree on the set of received events with high probability and deliver these in total order to the application. The intuition behind EpTO is to make events available in all processes quickly with high probability.
Once events are thought to be available in every peer, each peer deterministically orders them using the timestamp of each event and breaking ties with the id of the broadcaster peer, delivers them to the application accordingly.
\par
The main insight behind EpTO dissemination protocol is a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}. The \textit{Balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin, completely random and independent from other balls.
\par
Using the balls-and-bins approach we model processes as bins and events (or set of events) as balls. Based on this model we will calculate how many balls need to be \textit{thrown} such that each bin contains at least one ball with arbitrarily high probability. Using this approach the number of messages transmitted per process per round is logarithmic in the number of processes, therefore the number of messages sent in the system is low and uniform over all processes. Using these approaches, EpTO becomes highly scalable and resilient, having in mind that it also provides total order.

@@

\subsection{Total Order Algorithms}
A totally ordered set of events is a partially ordered set which satisfies one additional property:
\begin{itemize}
	\item Totality (trichotomy law): For any $a, b \in S$, either $a \leq b$  or $b \leq a$.
\end{itemize}
\todo[inline]{Explain advantages and usefulness of both systems}
\par
In other words, total order is an ordering that defines the exact order of every event in the system. On the other hand, partial ordering only defines the order between certain key events that depend on each other. Partial order can be useful since it is less costly to implement but, in some cases where the order of all events is important, for example in order to have an exact image of the state of the system at all time, we have to use total order.
